---
title: "halloween_sweets_analysis"
author: "Kerr McIntosh"
date: "21/07/2020"
output: html_document
---

**Project Summary**
I have compiled 3 separate years worth of data relating to a survey carried out around halloween 2015, 2016 and 2017 with the intent of finding out nostalgic feelings related to historic popular candy sweets gifted at halloween.  A  lot of the data in the sheet wasn't directly relevant to the sweets themselves ( and which our interest is in) and so I discarded such unrelated data from my analysis.

```{r}
library(tidyverse)
```

```{r}
halloween_clean_data <- read_csv(here::here("clean_data/halloween_candy_clean_data.csv"))
```


**1) There were 752,978 candy ratings carried out over the three years**
```{r}
num_candy_ratings <- halloween_clean_data %>%
  drop_na(reaction) %>% 
  nrow()
num_candy_ratings
```

**2) Average age of tricker treaters was 35 and non tricker treaters 39.**
```{r}
num_candy_ratings <- halloween_clean_data %>%
  select(survey_id, age, going_out) %>%
  group_by(going_out) %>%
  drop_na() %>% 
  summarise(average_age = round(mean(age, na.rm=TRUE)))

num_candy_ratings
```

**3) The halloween treat resulting in the most despair was a broken_glow_stick.  The treat resulting in the most joy was any full sized candy bar ( and the next best a brand reeses peanut butter cups).  The most indifference (Mehs) came from lollipops.**
```{r}
most_rated <- halloween_clean_data %>%
  group_by(reaction, treat) %>% 
  summarise(count = n()) %>% 
  filter(count == max(count))

most_rated           

```

**4)1,990 people rated Starburst with despair**
```{r}
starburst_despair <- halloween_clean_data %>%
  filter(str_detect(treat, "starburst")) %>% 
  filter(reaction == "despair" ) %>% 
  distinct(survey_id) %>% 
  nrow()
starburst_despair
```

```{r}
treat_scoring <- halloween_clean_data %>%
  mutate(score = case_when(
  reaction == "despair" ~ -1,
  reaction == "joy" ~ 1,
  reaction == "meh" ~ 0
))


treat_scoring
```

**6) the highest scoring candy bar for males was any full sized candy bar ( the highest brand was hershey's milk chocolate).  For females it was hershey's milk chocolate.**

```{r}
highest_gender <- treat_scoring %>% 
 select(gender, treat, score) %>%
  group_by(gender, treat) %>% 
  summarise(score = n()) %>% 
  slice(which.max(score)) %>% 
  drop_na() %>% 
  arrange(desc(score))
highest_gender

# detailed look at gender choices
highest_female <- treat_scoring %>% 
  filter(gender == "female") %>% 
  group_by(treat) %>% 
  summarise(score = n()) %>% 
  arrange(desc(score))
head(highest_female, 6)

highest_male <- treat_scoring %>% 
  filter(gender == "male") %>% 
  group_by(treat) %>% 
  summarise(score = n()) %>% 
  arrange(desc(score))
head(highest_male, 6)


```

**7) In 2015 the most popular bar was Reese's peanut butter cups. In 2016 it was Candy Corn or any full sized candy bar - both same score.  In 2017 it was peanut butter MnMs.**
```{r}
treat_scoring <- treat_scoring %>% 
  mutate(year = case_when(
  str_detect(survey_id, "2015") ~ "2015",
  str_detect(survey_id, "2016") ~ "2016",
  str_detect(survey_id, "2017") ~ "2017",
))

highest <- treat_scoring %>% 
  group_by(year, treat)  %>% 
  summarise(score = n()) %>% 
  filter(score== max(score))
highest


```

**Below is a list of the most popular candy bars in each surveyed country**
```{r}
treat_by_country <- treat_scoring %>% 
  select(country, treat, score) %>% 
  group_by(country, treat) %>% 
  summarise(most_popular = n()) %>% 
  slice(which.max(most_popular)) %>% 
  drop_na() %>% 
  arrange(desc(most_popular))

treat_by_country 
  
```
 
